Simulation Modeling of Glycosylation Effects on Potassium Channels of Mouse Cardiomyocytes

### 2021-06-22 draft_plos.tex
References added

Thus, we investigated the impact of reduced hybrid/complex N-glycosylation through the lens of electrophysiology. Electrophysiology refers to studies of electrical properties in scales from ion channels in cells to organ systems \cite{scanziani2009electrophysiology}. Electrical signaling in the heart has vital functions related to intra and extracellular communication, rhythmicity of heartbeats, and provides a driving force for contraction. The action potential (AP) is the net transmembrane potential varying over time which results from the composite activities of different ion channels \cite{grant2009cardiac}. Fig~\ref{fig1}-(a) shows the simulated AP of a mouse ventricular apex myocyte using models adopted from \cite{bondarenko2004computer}. The AP is the result of orchestrated activities of various ionic currents described in Fig-(b). Even small changes in VGIC function can contribute to aberrant AP waveform and/or conduction, leading to arrhythmias. As illustrated in Fig~\ref{fig2}, Na\textsuperscript{+} channels (Na\textsubscript{v}) are responsible for AP initiation, or the depolarization phase, Ca\textsuperscript{2+} channels (Ca\textsubscript{v}) drive the prolonged depolarization phase particularly in larger species (“plateau”), while several K\textsuperscript{+} channel (K\textsubscript{v}) isoforms are collectively responsible for AP deactivation, i.e., repolarization. Like most transmembrane proteins, VGICs are heavily glycosylated membrane proteins \cite{ednie2011modulation}. Studies have proven that glycosylation can affect VGIC function predominantly. For example, we reported that a saturating, electrostatic effect of negatively charged sialic acids, which are typically attached to the terminal of glycan branches, significantly altered cardiac electrical signaling \cite{ednie2013expression, ednie2015reduced}. Our recent studies observed aberrant electrical signaling, e.g., prolonged APs and abnormal early re-activations, in Mgat1KO ventricular apex myocytes \cite{ednie2019reduced, ednie2019reduced2}. These observations strongly support direct and disease-related effects of reduced N-glycosylation on VGIC gating and activity.

@@

Although these in-vitro experiments showed changes in activities of ion channels, there are certain limitations that include: 1) \textit{Detailed ion channel kinetics are difficult to determine rigorously using in-vitro experiments alone}. From the electrophysiological experiment, pathological electrical signaling can be observed through measurements of APs at the whole-cell level and currents at the ion-channel level. However, it is difficult to relate, rigorously, the impacts of VGIC gating changes to the altered AP waveform/conduction and vice versa. 2) \textit{Segregating different K\textsuperscript{+} currents (I\textsubscript{K}) is difficult using whole-cell recording experimental techniques only}. I\textsubscript{Ksum} is the result of joint activity of multiple isoforms with each isoform producing a different, but slightly overlapping (in activation and inactivation voltages) component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) \cite{du2017} that contributes to various portions of AP repolarization. However, in-vitro experiments can only measure the sum of these component currents (I\textsubscript{Ksum}) using whole-cell voltage-clamp protocols. Even pharmacologic separation of K\textsubscript{v} current types is difficult, as the specificity of drugs for a single current type is not ideal. Hence, component K\textsuperscript{+} currents are usually estimated through multiple exponential fits of the decaying portion (due to channel inactivation) of the total K\textsuperscript{+} current \cite{brunet2004heterogeneous}. Although the exponential function illustrates the shape of a component-current trace well, it cannot adequately describe kinetic dynamics of the current, nor fully and rigorously distinguish among currents produced by different K\textsubscript{v} isoforms because of their slightly overlapping voltage-dependence of gating. As K\textsubscript{v} have a critical function in repolarizing the cell, it is imperative to model and examine K\textsubscript{v} isoforms thoroughly.

@@

\section*{Background}
% === II. RESEARCH BACKGROUND ===========================================
% =======================================================================
\subsection*{Glycosylation and DCM}
Protein glycosylation is an essential cellular process that impacts many cell functions \cite{marques2017cardiac}. Briefly, protein glycosylation is the sequential co-/post-translational process of attaching sugar residues (glycans) to proteins. Cardiac VGICs are heavily glycosylated proteins with upwards of 30\% of their mass consisting of N- and O-linked glycans. A growing number of cardiac diseases, including DCM and hypertrophic cardiomyopathy, can present with concurrent, albeit, modest changes in glycosylation \cite{gehrmann2003cardiomyopathy, footitt2009cardiomyopathy, marques2017cardiac}. Mgat1 expression was implicated in cardiac function and shown to be down-regulated in human end-stage idiopathic DCM. Genome-wide searches identified changes in glycosylation-related gene expression in human idiopathic DCM, including glycosyltransferases \cite{barrans2002global, hwang2002microarray, yung2004gene}; and proteomic/glycomic studies show changes in serum N-glycosylation in heart disease models and in humans with DCM risk factors \cite{nishio1995identification, knezevic2009variability, miura2016glycomics, nagai2016aberrant, yang2015glycoproteins}. Models of DCM/heart failure were associated with subtle changes in glycosylation of proteins involved in electromechanical processes, and ~20\% of patients with congenital disorders of glycosylation (CDG) present with cardiac deficits, including idiopathic DCM \cite{gehrmann2003cardiomyopathy, marques2017cardiac}. The data suggest a correlation between modest changes in extracellularly facing glycosylation and DCM/heart disease.

\subsection*{Simulation Modeling of Mouse Ventricular Myocytes}
Electrophysiology was pioneered by Hodgkin and Huxley with a series of well-known experiments on the squid giant axon \cite{hodgkin1952quantitative}. Their simulation model was built upon two fundamental principles observed in the experimental data. 1) Cells are selectively activated by different ion channels. 2) Activities of ion channels to produce ionic currents are controlled through voltage changes, from which the name ``voltage-gated ion channels'' originates. The voltage-gating mechanism is encoded in the simulations model with partial differential equations. Not only are simulation models compatible with data and descriptive, but they also simulate changes of AP waveforms and ion channel activity. Thus, models provide connections among ionic currents, action potentials, and disease effects. However, significant domain knowledge and in-vitro investigations are required, because simulation models often have complex structures based on nonlinear functions and differential equations with a high-dimensional set of parameters.

@@

\begin{thebibliography}{10}
    \bibitem{cdc2019deaths}
    Heron M.
    \newblock Deaths: Leading Causes for 2017.
    \newblock National Vital Statistics Reports. 2019;68(6).

@@

    \bibitem{scanziani2009electrophysiology}
    Scanziani M, H{\"a}usser M.
    \newblock Electrophysiology in the age of light.
    \newblock Nature. 2009;461(7266):930--939.

@@

@@

@@

@@

    \bibitem{schwetz2010n}
    Schwetz TA, Norring SA, Bennett ES.
    \newblock N-glycans modulate Kv1. 5 gating but have no effect on Kv1. 4 gating.
    \newblock Biochimica et Biophysica Acta (BBA)-Biomembranes.
    2010;1798(3):367--375.
\end{thebibliography}

>>>>

Thus, we investigated the impact of reduced hybrid/complex N-glycosylation through the lens of electrophysiology. Electrophysiology refers to studies of electrical properties in scales from ion channels in cells to organ systems. Electrical signaling in the heart has vital functions related to intra and extracellular communication, rhythmicity of heartbeats, and provides a driving force for contraction \cite{koenig2011voltage}. The action potential (AP) is the net transmembrane potential varying over time which results from the composite activities of different ion channels \cite{grant2009cardiac}. Fig~\ref{fig1}-(a) shows the simulated AP of a mouse ventricular apex myocyte using models adopted from \cite{bondarenko2004computer}. The AP is the result of orchestrated activities of various ionic currents described in Fig-(b). Even small changes in VGIC function can contribute to aberrant AP waveform and/or conduction, leading to arrhythmias. As illustrated in Fig~\ref{fig2}, Na\textsuperscript{+} channels (Na\textsubscript{v}) are responsible for AP initiation, or the depolarization phase, Ca\textsuperscript{2+} channels (Ca\textsubscript{v}) drive the prolonged depolarization phase particularly in larger species (“plateau”), while several K\textsuperscript{+} channel (K\textsubscript{v}) isoforms are collectively responsible for AP deactivation, i.e., repolarization. Like most transmembrane proteins, VGICs are heavily glycosylated membrane proteins \cite{ednie2011modulation}. Studies have proven that glycosylation can affect VGIC function predominantly. For example, we reported that a saturating, electrostatic effect of negatively charged sialic acids, which are typically attached to the terminal of glycan branches, significantly altered cardiac electrical signaling \cite{ednie2013expression, ednie2015reduced}. Our recent studies observed aberrant electrical signaling, e.g., prolonged APs and abnormal early re-activations, in Mgat1KO ventricular apex myocytes \cite{ednie2019reduced, ednie2019reduced2}. These observations strongly support direct and disease-related effects of reduced N-glycosylation on VGIC gating and activity.

@@

Although these in-vitro experiments showed changes in activities of ion channels, there are certain limitations that include: 1) \textit{Detailed ion channel kinetics are difficult to determine rigorously using in-vitro experiments alone}. From the electrophysiological experiment, pathological electrical signaling can be observed through measurements of APs at the whole-cell level and currents at the ion-channel level. However, it is difficult to relate, rigorously, the impacts of VGIC gating changes to the altered AP waveform/conduction and vice versa. 2) \textit{Segregating different K\textsuperscript{+} currents (I\textsubscript{K}) is difficult using whole-cell recording experimental techniques only}. I\textsubscript{Ksum} is the result of joint activity of multiple isoforms with each isoform producing a different, but slightly overlapping (in activation and inactivation voltages) component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) that contributes to various portions of AP repolarization \cite{du2017, jerng2021light}. However, in-vitro experiments can only measure the sum of these component currents (I\textsubscript{Ksum}) using whole-cell voltage-clamp protocols. Even pharmacologic separation of K\textsubscript{v} current types is difficult, as the specificity of drugs for a single current type is not ideal. Hence, component K\textsuperscript{+} currents are usually estimated through multiple exponential fits of the decaying portion (due to channel inactivation) of the total K\textsuperscript{+} current \cite{brunet2004heterogeneous}. Although the exponential function illustrates the shape of a component-current trace well, it cannot adequately describe kinetic dynamics of the current, nor fully and rigorously distinguish among currents produced by different K\textsubscript{v} isoforms because of their slightly overlapping voltage-dependence of gating. As K\textsubscript{v} have a critical function in repolarizing the cell, it is imperative to model and examine K\textsubscript{v} isoforms thoroughly.

@@

\section*{Background}
% === II. RESEARCH BACKGROUND ===========================================
% =======================================================================
\subsection*{Glycosylation and DCM}
Protein glycosylation is an essential cellular process that impacts many cell functions \cite{marques2017cardiac}. Briefly, protein glycosylation is the sequential co-/post-translational process of attaching sugar residues (glycans) to proteins. Cardiac VGICs are heavily glycosylated proteins with upwards of 30\% of their mass consisting of N- and O-linked glycans. A growing number of cardiac diseases, including DCM and hypertrophic cardiomyopathy, can present with concurrent, albeit, modest changes in glycosylation \cite{gehrmann2003cardiomyopathy, footitt2009cardiomyopathy, molina2013differential, marques2017cardiac}. Mgat1 expression was implicated in cardiac function and shown to be down-regulated in human end-stage idiopathic DCM. Genome-wide searches identified changes in glycosylation-related gene expression in human idiopathic DCM, including glycosyltransferases \cite{barrans2002global, hwang2002microarray, yung2004gene}; and proteomic/glycomic studies show changes in serum N-glycosylation in heart disease models and in humans with DCM risk factors \cite{nishio1995identification, knezevic2009variability, miura2016glycomics, nagai2016aberrant, yang2015glycoproteins}. Models of DCM/heart failure were associated with subtle changes in glycosylation of proteins involved in electromechanical processes, and ~20\% of patients with congenital disorders of glycosylation (CDG) present with cardiac deficits, including idiopathic DCM \cite{gehrmann2003cardiomyopathy, marques2017cardiac}. The data suggest a correlation between modest changes in extracellularly facing glycosylation and DCM/heart disease.

\subsection*{Simulation Modeling of Mouse Ventricular Myocytes}
Electrophysiology was pioneered by Hodgkin and Huxley with a series of well-known experiments on the squid giant axon \cite{hodgkin1952quantitative}. Their simulation model was built upon two fundamental principles observed in the experimental data. 1) Cells are selectively activated by different ion channels. 2) Activities of ion channels to produce ionic currents are controlled through voltage changes, from which the name ``voltage-gated ion channels'' originates. The voltage-gating mechanism is encoded in the simulations model with partial differential equations. Not only are simulation models compatible with data and descriptive, but they also simulate changes of AP waveforms and ion channel activity \cite{mullins2013mathematical}. Thus, models provide connections among ionic currents, action potentials, and disease effects. However, significant domain knowledge and in-vitro investigations are required, because simulation models often have complex structures based on nonlinear functions and differential equations with a high-dimensional set of parameters.

@@

% \bibliography{ref}

\begin{thebibliography}{10}

\bibitem{cdc2019deaths}
Heron M.
\newblock Deaths: Leading Causes for 2017.
\newblock National Vital Statistics Reports. 2019;68(6).

@@

\bibitem{koenig2011voltage}
Koenig X, Dysek S, Kimbacher S, Mike AK, Cervenka R, Lukacs P, et~al.
\newblock Voltage-gated ion channel dysfunction precedes cardiomyopathy
  development in the dystrophic heart.
\newblock PLoS One. 2011;6(5):e20300.

@@

\bibitem{jerng2021light}
Jerng HH, Patel JM, Khan TA, Arenkiel BR, Pfaffinger PJ.
\newblock Light-regulated voltage-gated potassium channels for acute
  interrogation of channel function in neurons and behavior.
\newblock Plos one. 2021;16(3):e0248688.

@@

\bibitem{molina2013differential}
Molina-Navarro MM, Rosell{\'o}-Llet{\'\i} E, Ortega A, Taraz{\'o}n E, Otero M,
  Mart{\'\i}nez-Dolz L, et~al.
\newblock Differential gene expression of cardiac ion channels in human dilated
  cardiomyopathy.
\newblock PloS one. 2013;8(12):e79792.

@@

\bibitem{mullins2013mathematical}
Mullins PD, Bondarenko VE.
\newblock A mathematical model of the mouse ventricular myocyte contraction.
\newblock PLoS One. 2013;8(5):e63141.

@@

\bibitem{schwetz2010n}
Schwetz TA, Norring SA, Bennett ES.
\newblock N-glycans modulate Kv1. 5 gating but have no effect on Kv1. 4 gating.
\newblock Biochimica et Biophysica Acta (BBA)-Biomembranes.
  2010;1798(3):367--375.

\end{thebibliography}
